Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-12 vesicle_code/docs/permeableVes/main.tex
yny

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Ashley Gannon}
\author[b]{Yuan-Nan Young} 
\author[a,1]{Bryan Quaife}

@@

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally.}
%\authordeclaration{Please declare any competing interests here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{1} E-mail: bquaife\@fsu.edu}

@@

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}.  Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
Water transport in and out of cell is also found to be important to 
active cell shape fluctuations leading to bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.

The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. Without aquaporins or other membrane channels that promotes water permeability, the intrinsic water permeability is known to depend on the temperature \cite{}, solvent ph \cite{}, and the hydration  energy of the membrane \cite{}. In the red blood cell membrane, aquaporin-1 (AQP1) channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB}.

In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient across the membrane to neutralize the osmolarity.   The resultant swelling/shrinking of cell is inevitably coupled to changes in membrane tension and cell shape \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL}.
In this work we focus on the intrinsic water permeability and study the subtle balance between hydrodynamics and elasticity of a semipermeable membrane under various flowing conditions and confinement. Specifically we focus on the effects of intrinsic water permeability in a vesicle under stress due to flows and/or confinement.

Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. In these situations, the osmotic stress is the dominant driving force for water transport across the membrane. Thus most modeling works ignore the mechanical effects (capillary pressure, bending force and membrane tension) on vesicles under an osmotic shock. However, the osmotic water flux is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.

\paragraph{Literature to maybe cite}
\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
happens in regions where proteins congregate and this can lead to
semi-permeability?

\section*{Governing equations \label{sc:formulate}}
We consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid driven by the imposed flow $\uu_\infty(\xx)$.
Since fluid is transported across the semi-permeable vesicle membrane,
we only consider fluids with uniform viscosity. Letting $\gamma$ be the
vesicle interface, the fluid velocity $\uu$ and pressure $p$ satisfy
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, \quad
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension is $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Numerical Methods}

@@

>>>>

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Bryan Quaife}
\author[a]{Ashley Gannon}
\author[b,1]{Y.-N. Young}

@@

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally.}
%\authordeclaration{Please declare any competing interests here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
%\correspondingauthor{\textsuperscript{1} E-mail: bquaife\@fsu.edu}
\correspondingauthor{\textsuperscript{1} E-mail: yyoung@njit.edu}

@@

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}: Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
Cells also need to control water transport across the cell membrane as they activate shape fluctuations prior to the bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.

The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. Without aquaporins or other membrane channels that promotes water permeability, the intrinsic water permeability is known to depend on the temperature \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, lipid composition \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, and the hydration  energy of the membrane \cite{MarrinkBerendsen1994_JPhysChem}. In the red blood cell membrane, aquaporin-1 channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB,BhatiaRobinsonDimova2020_SoftMatt}.

The swelling/shrinking of a cell is inevitably coupled to changes in membrane tension and cell shape \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL}.  
Most studies focus on the osmolarity-induced water flux, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
In this work we focus on the intrinsic water permeability and study the subtle balance between hydrodynamics and elasticity of a semipermeable membrane under various flowing conditions and confinement. Specifically we focus on the effects of intrinsic water permeability in a vesicle under stress due to flows and/or confinement.

Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB,BhatiaChrist2020_SoftMatt}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. When trapped in a microfluidic channel, an osmotic shock induces 
 vesicle shape changes, tank-treading and tumbling motions \cite{BhatiaRobinsonDimova2020_SoftMatt,BhatiaChrist2020_SoftMatt}, an illustration that the permeable water flow is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.  In this work we use numerical simulations to examine the hydrodynamics of a semipermeable vesicle in two dimensions. Focusing on the effects of water permeability on the hydrodynamics of a vesicle under mechanical stresses, we investigate how a semipermeable vesicle may behave differently from their impermeable counterpart in familiar situations such as shear flows, a stenosis channel, and a contracting channel.

%
%\paragraph{Literature to maybe cite}
%\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
%happens in regions where proteins congregate and this can lead to
%semi-permeability?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Formulation}

@@

\section*{Supplementary Material: Formulation \label{sc:formulate}}
We consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid driven by the imposed flow $\uu_\infty(\xx)$.
Since fluid is transported across the semi-permeable vesicle membrane,
we only consider fluids with uniform viscosity. Letting $\gamma$ be the
vesicle interface, the fluid velocity $\uu$ and pressure $p$ satisfy
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, \quad
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension is $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Numerical Methods}
